---
title: "logicals"
author: "Mia Forsline"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### Taking a look at `starwars`
- `include =`: show or hide everything
- `echo =`: show or hide code
```{r, include = FALSE}
#show nothing from this code chunk (include = FALSE)
#show only outputs not code ()

head(starwars)

dim(starwars)

summary(starwars)
```

- `warning =`: remove warnings
- `eval =`: show code w/out running it and showing the output 
```{r, warning = FALSE}
#remove warnings of missing data in the plot 
ggplot(data = starwars, aes(x = height, y = mass))+
  geom_point(color = "dark blue")+
  theme_classic()
```
### Examine individual columns - find the mean, variance, & SD of character height 
```{r}
sw_height_mean <- mean(starwars$height, na.rm = TRUE)
sw_height_mean
#mean function will not remove NA values automatically (na.rm = FALSE) - we must change it manually (na.rm = TRUE)
#original data only measures up to nearest cm, so our mean should be reported to the nearest tenth of a cm or so 

#find the variance of character height 
#find standard deviation of character height
sw_height_var <- var(starwars$height, na.rm = TRUE)
sw_height_var #cm^2 units
sw_height_sd <- sd(starwars$height, na.rm = TRUE)
sw_height_sd #cm units 
```

Mean character height is `r round(sw_height_mean, 1)` cm, with a standard deviation of `r round(sw_height_var, 1)`cm. 

## Logical expressions
```{r}
5 < 10

#alphabetical
"cat" > "armadillo"
"cat" > "elephant"
"dog" == "dog"
```

```{r}
10 <= 11
5 >= 4
3 <=1
```

```{r}
tarantulas <- c(1, 3, 7, 10, 15)

tarantulas > 5
```

## Logical operators:
- `!`
- `&`
- `|`
```{r}
3 <1 | 8 > 2

6 < 10 & 5 < 1 

7 != 11 

4 != 4 | 2 < 5 | 6 > 10

#R sees 0 as false and 1 as true 
0 == FALSE
1 == TRUE
```

